\section{Problem Description}
\label{sec:problem}

Bubbles, x-axis -> income, y-axis -> life expectation, size -> population total, color -> continent, label -> nation.

\subsection{Questions}
Gapminder tried to answer to many questions.

\paragraph{Concrete Questions}
\begin{itemize}

    % population size
    \item Where are the countries with the largest population located?
    \item Which continent has the largest population?

    % life expectation
    \item What is the life expectation of people living in the Netherlands in 2015?    
    \item What is the life expectation of the world population?
    \item Which countries have the highest life expectation?
    \item What was the life expectation of people living in Germany during the Second World War?
    \item How has the life expectation of people living in Italy evolved from 1800 to today?    
    \item 

    % income
    \item What is the average world income income per person?
    \item Which countries have the highest income?
    \item Which countries have the lowest income?
    \item What is the average income of the population of the United States?
    \item Are the average incomes of small countries above or under the world average?

    % correlation
    \item Do the countries with the highest income also have the highest life expectation?
    \item Do countries closed to each other have similar incomes?
    \item Do countries closed to each other have similar life expectations?

\end{itemize}

\paragraph{Higher-level Questions}
\begin{itemize}

    % life expectation
    \item What is the global trend for the life expectation?

    % correlation
    \item How are life expectation and income correlated to each other today?    
    \item Did the correlation between life expectation and income change over time?
    \item Does the geographical area influence the correlation between life expectation and income?
    \item Does the size of a population influence the life expectation?
    
\end{itemize}

\paragraph{Summary}
\begin{itemize}    
    \item What is the world situation in terms of population, average income, life expectation?
    \item How did the situation change over the last 2 centuries?
    \item What is the trend for the next 10 years?
\end{itemize}

\subsection{Users}
Each dataset in Gapminder has time and geographical information.
Datasets represents all kind of indicators from many different sectors.
The range of users for this tool is big.
Typical users can be politicians, analysts, journalists, teachers, researchers.

\subsection{Purpose}
The purpose of the visualization is to present a picture of the global world situation from many different points of view.
The tool allows to easily compare many different indicators, analyze the situation of different countries in the world, study the historical evolution of the global world situation and predict trends for the next years.
